---
title: "Reproduce Reports"
output: html_document
params: 
  continent: Americas
bibliography: citations.bib
---

Chunks named "setup" are a little special. They will always run first in R Notebooks and interactively if they haven't been run yet.

```{r setup, echo = FALSE, message = FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse)
library(gapminder)
```

## Your turn 1

Create a new R Markdown file. Go to File > New File > R Markdown. Press OK. Save the file and press the "Knit" button above.

## Your turn 2

Do the ten-twenty minute tutorial on markdown at https://commonmark.org/help/tutorial. Let us know if you need help!

## Your turn 3

1. Use Markdown syntax to stylize the text from the [Gapminder website](https://www.gapminder.org/data/documentation/gd001/) below. Experiment with bolding, italicizing, making lists, etc.

GDP per capita measures the value of everything produced in a country during a year, divided by the number of people. The unit is in international dollars, fixed 2011 prices. The data is adjusted for inflation and differences in the cost of living between countries, so-called PPP dollars. The end of the time series, between 1990 and 2016, uses the latest GDP per capita data from the World Bank, from their World Development Indicators. To go back in time before the World Bank series starts in 1990, we have used several sources, such as Angus Maddison. 

## Your turn 4

Create a code chunk. You can type it in manually, use the keyboard short-cut (Cmd/Ctrl + Option/Alt + I), or use the "Insert" button above. Put the following code in it:

gapminder %>% 
  slice(1:5) %>% 
  knitr::kable()
  
Knit the document

## Your turn 5
  
Add `echo = FALSE` to the code chunk above and re-knit

Remove `echo = FALSE` from the code chunk and move it to `knitr::opts_chunk$set()` in the `setup` code chunk. Re-knit. What's different about this? Look around the rest of the document.

When you've figured it out, remove `echo = FALSE` from `knitr::opts_chunk$set()`.

## Your turn 6

1. Remove `eval = FALSE` so that R Markdown evaluates the code.
2. Use `summarize()` and `n_distinct()` to get the the number of unique years in gapminder and save the results as `n_years`.
3. Use inline code to describe the data set in the text below the code chunk and re-knit.

```{r, eval = FALSE}
___ ___ gapminder %>% 
  _________ %>% 
  # pull() extracts a column as a vector
  # if no column is called, it uses the most recently created
  pull()
```

Gapminder observed data over _________ different years.


## Your turn 7

1. Set figure chunk options such as `dpi`, `fig.width`, and `fig.height`. Run `knitr::opts_chunk$get()` in the console to see the defaults.
2. Change the YAML header above from `output: html_document` to another output type like `pdf_document` or `word_document`.
3. Add your name to the YAML header using `author: Your Name`.

```{r}
gapminder %>% 
  filter(continent != "Oceania") %>% 
  ggplot(aes(x = year, y = lifeExp, group = country, color = country)) +
  geom_line(lwd = 1, show.legend = FALSE) + 
  facet_wrap(~ continent) +
  # `country_colors` is an object included in gapminder
  scale_color_manual(values = country_colors) + 
  theme_minimal(14) + 
  theme(strip.text = element_text(size = rel(1.1)))
```


## Your turn 8

1. Change the `params` option in the YAML header to use a different continent. Re-knit

```{r}
gapminder %>% 
  filter(continent == params$continent) %>% 
  ggplot(aes(x = year, y = lifeExp, group = country, color = country)) +
  geom_line(lwd = 1, show.legend = FALSE) + 
  scale_color_manual(values = country_colors) +
  theme_minimal(14) + 
  theme(strip.text = element_text(size = rel(1.1))) +
  ggtitle(paste("Continent:", params$continent))
```

## Your turn 9

1. Cite the Causal Inference book in text below in the format ```[@citation-label]``` (without the tick marks; you only need those to tell R Markdown not to evalaute the this). The label for the citation is `hernan_causal_2019`
2. Add the American Journal of Epidemiology CSL to the YAML using `
csl: aje.csl`

We calculated the average treatment effect using g-computation ________.

# Take aways

* R Markdown has simple syntax that's easy to use
* Your analysis and narrative are easy to intertwine in R Markdown
* You can re-run your analysis with different inputs and outputs
* Code chunks provide a variety of options to change the resulting report
* Use YAML to set options for R Markdown
* R Markdown supports a powerful citation system. Find more CSL styles at https://github.com/citation-style-language/styles


