---
title: "Quarto: Whole Game"
format: html
---

```{r}
#| label: setup
library(tidyverse)
# this loads the `copy_qmd_template()` helper function we'll use in this exercise
source("copy_qmd_template.R")
```

*For this set of exercises, we'll work through a Quarto report in `whole_game.qmd`. Open up that file and follow along with these exercises:*

## *Your Turn 1*

1.  *Turn on Visual Quarto mode for `exercises.qmd` (this document) AND `whole_game.qmd`*
2.  *In your local R session, run the code in the code chunk called "prep-data" in `whole_game.qmd`. Look at the chunk options (everything after at the top that starts with `#|`. What do you think setting the `include` option does?*
3.  *Render the document and take a look. Were you right?*

## *Your Turn 2*

1.  *Add a citation to the methods section for the data. Use the citation wizard to search PubMed for an article called "Prevalence of coronary heart disease risk factors among rural blacks: A community-based study"*

## *Your Turn 3*

*Something not quite working from the last exercise? Run this code to update `whole_game.qmd` and catch up*

```{r}
#| eval: false
#| echo: false
copy_qmd_template("your_turn_3.qmd")
```

1.  *Take a look at the chunk called "descriptive-stats." What do you think `echo: false` does, and how is it different from `include: false`? What do you think `eval: false` means?*
2.  *Read the code in "descriptive-stats". What are we preparing?*
3.  *Look at the narrative text right after "descriptive-stats". It's full of bits of code. What are we doing here, and what do you think the output should look like?*
4.  *Change `eval` to `true` in the "descriptive-stats" code chunk and re-render. Look at this part of the document. What does the output look like?*

## *Your Turn 4*

*Something not quite working from the last exercise? Run this code to update `whole_game.qmd` and catch up*

```{r}
#| eval: false
#| echo: false
copy_qmd_template("your_turn_4.qmd")
```

1.  *In the code chunk called "tbl-one", create a descriptive table by filling in the first blank with `tbl_summary()`* from the gtsummary package, a great tool for making tables.
2.  *Pipe the results of `tbl_summary()` to `add_difference()`* (put this function in the second blank)
3.  Read through the code and make a prediction about what you'll see. What do you think `tbl-cap`, at the top of the code chunk, will do? What do you think `by = diabetic` in `tbl_summary()` will do?
4.  *Change `eval` to `true` and re-render.*

## *Your Turn 5*

*Something not quite working from the last exercise? Run this code to update `whole_game.qmd` and catch up*

```{r}
#| eval: false
#| echo: false
copy_qmd_template("your_turn_5.qmd")
```

1.  *In the code chunk called "fig-one", load the patchwork library*

2.  *At the end of the code chunk, include this code: `fig1a + fig1b` . (Here, we're using the patchwork library, which allows you to join ggplots side by side with `+`).*

3.  *In "fig-one", add a chunk option: `#| fig-cap: "A1c and diabetes status by WH ratio"` . Then, where it says "\_\_\_\_\_", add* \@fig-one

4.  *In "fig-one", change `eval` to `true` and re-render*

## *Your Turn 6*

*Something not quite working from the last exercise? Run this code to update `whole_game.qmd` and catch up*

```{r}
#| eval: false
#| echo: false
copy_qmd_template("your_turn_6.qmd")
```

1.  *In the code chunk called "tbl-two", pass the results of both `lm()` and `glm()` to `tbl_regression()`*
2.  In the narrative text following the code chunk, fill in the blank with @tbl-two
3.  Take a look at the `tbl_merge()` function call. What do you think this code is doing to the to the two objects created by `tbl_regression()`? Predict what the result will be.
4.  *change `eval` to `true` and re-render*

***Stretch goal**: The description for Table 2 is pretty sparse. It would probably be better to have some numbers included in the narrative text. Read gtsummary's [vignette on inline text](http://www.danieldsjoberg.com/gtsummary/articles/inline_text.html) and add a few important details.*

## *Your Turn 7*

*Something not quite working from the last exercise? Run this code to update `whole_game.qmd` and catch up*

```{r}
#| eval: false
#| echo: false
copy_qmd_template("your_turn_7.qmd")
```

*Previously, we only had data for one location in the study---Buckingham. We've just received data from the other site, Louisa, and we need to update the tables, figures, and narrative text...*

*...but since we used Quarto, we can just load the new data set, and R will update everything for us.*

1.  *Change `read_csv()` to load `diabetes.csv` instead of `diabetes-buckingham.csv`*
2.  *Add the `location` variable to Table 1*
3.  *re-render. How's it look?*
