---
title: "Reproduce Reports"
output: html_document
params: 
  continent: Americas
bibliography: citations.bib
---

Chunks named "setup" are a little special.
They will always run first in R Notebooks and interactively if they haven't been run yet.

```{r setup, echo = FALSE, message = FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse)
library(gapminder)

copy_rmd_template <- function(template) {
  copied <- file.copy(
    file.path("templates", template),
    "exercises.Rmd",
    overwrite = TRUE
  )
  
  if (!copied) {
    stop(
      file.path("templates", template), " not found. ", 
      "Are you sure you're in the right working directory?",
      call. = FALSE
    )
  }
  
  template_r <- usethis::ui_path(file.path("templates", template))
  targets_r <- usethis::ui_path("exercises.Rmd")
  usethis::ui_done("Copying {template_r} to {targets_r}")
}
```

## Your turn 1

Create a new R Markdown file.
Go to File \> New File \> R Markdown.
Press OK. Save the file and press the "Knit" button above.

## Your turn 2

Read this short introduction to Visual R Markdown: <https://rstudio.github.io/visual-markdown-editing/#/intro?id=getting-started>

Turn on Visual R Markdown mode for this document.
Then, stylize the text from the [Gapminder website](https://www.gapminder.org/data/documentation/gd001/) below.
Experiment with bolding, italicizing, making lists, etc.

GDP per capita measures the value of everything produced in a country during a year, divided by the number of people.
The unit is in international dollars, fixed 2011 prices.
The data is adjusted for inflation and differences in the cost of living between countries, so-called PPP dollars.
The end of the time series, between 1990 and 2016, uses the latest GDP per capita data from the World Bank, from their World Development Indicators.
To go back in time before the World Bank series starts in 1990, we have used several sources, such as Angus Maddison.

## Your turn 3

Create a code chunk.
You can type it in manually, use the keyboard short-cut (Cmd/Ctrl + Option/Alt + I), or use the "Insert" button above.
Put the following code in it:

```{r}
gapminder %>% slice(1:5) %>% knitr::kable()
```

Knit the document

## Your turn 4

Something not quite working from the last exercise?
Run this code to update `exercises.Rmd` and catch up

```{r, eval=FALSE, echo = FALSE}
copy_rmd_template("your_turn_4.Rmd")
```

Add `echo = FALSE` to the code chunk above and re-knit

Remove `echo = FALSE` from the code chunk and move it to `knitr::opts_chunk$set()` in the `setup` code chunk.
Re-knit.
What's different about this?
Look around the rest of the document.

When you've figured it out, remove `echo = FALSE` from `knitr::opts_chunk$set()`.

## Your turn 5

Something not quite working from the last exercise?
Run this code to update `exercises.Rmd` and catch up:

```{r, eval=FALSE, echo = FALSE}
copy_rmd_template("your_turn_5.Rmd")
```

In the code chunk below:

1.  Remove `eval = FALSE` so that R Markdown evaluates the code.
2.  Use `summarize()` and `n_distinct()` to get the the number of unique years in gapminder and save the results as `n_years`.
3.  Use inline code to describe the data set in the text below the code chunk and re-knit.

```{r, eval = FALSE}
___ ___ gapminder %>% 
  _________ %>% 
  # pull() extracts a column as a vector
  # if no column is called, it uses the most recently created
  pull()
```

Gapminder observed data over \_\_\_\_\_\_\_\_\_ different years.

## Your turn 6

Something not quite working from the last exercise?
Run this code to update `exercises.Rmd` and catch up:

```{r, eval=FALSE, echo = FALSE}
copy_rmd_template("your_turn_6.Rmd")
```

In the code chunk below:

1.  Set figure chunk options to the code chunk below, such as `dpi`, `fig.width`, and `fig.height`.
    Run `knitr::opts_chunk$get()` in the console to see the defaults.

2.  Add your name to the YAML header using `author: Your Name`.

3.  Change the YAML header above from `output: html_document` to `output: distill::distill_article`.

4.  Set `distill::distill_article` to use the `toc` and `code_folding` options and re-knit (see `?distill::distill_article` for details on these and other arguments. Add them to the YAML like this (be mindful of the indents! Also, make sure you have Visual R Markdown turned on or the code below might look a little different):

    ``` {.yaml}
    output: 
      distill::distill_article:
        _________: true
        _________: true 
    ```

```{r}
gapminder %>% 
  filter(continent != "Oceania") %>% 
  ggplot(aes(x = year, y = lifeExp, group = country, color = country)) +
  geom_line(lwd = 1, show.legend = FALSE) + 
  facet_wrap(~ continent) +
  # `country_colors` is an object included in gapminder
  scale_color_manual(values = country_colors) + 
  theme_minimal(14) + 
  theme(strip.text = element_text(size = rel(1.1)))
```

## Your turn 7

Something not quite working from the last exercise?
Run this code to update `exercises.Rmd` and catch up:

```{r, eval=FALSE, echo = FALSE}
copy_rmd_template("your_turn_7.Rmd")
```

In the code chunk below:

1.  Take a look at the output of this plot in the document you've knitted.
2.  Now, change the `params` option in the YAML header to use a different continent.
3.  Re-knit. What's changed?

```{r}
gapminder %>% 
  filter(continent == params$continent) %>% 
  ggplot(aes(x = year, y = lifeExp, group = country, color = country)) +
  geom_line(lwd = 1, show.legend = FALSE) + 
  scale_color_manual(values = country_colors) +
  theme_minimal(14) + 
  theme(strip.text = element_text(size = rel(1.1))) +
  ggtitle(paste("Continent:", params$continent))
```

## Your turn 8

Something not quite working from the last exercise?
Run this code to update `exercises.Rmd` and catch up:

```{r, eval=FALSE, echo = FALSE}
copy_rmd_template("your_turn_8.Rmd")
```

In the sentence below:

1.  Cite the Causal Inference book in text below. Using the citation wizard, find the right citation under My sources \> Bibliography.
2.  Add the American Journal of Epidemiology CSL to the YAML using `csl: aje.csl`
3.  Re-knit

We calculated the average treatment effect using g-computation \_\_\_\_\_\_\_\_.

# Take aways

-   R Markdown has simple syntax that's easy to use
-   Your analysis and narrative are easy to intertwine in R Markdown
-   You can re-run your analysis with different inputs and outputs
-   Code chunks provide a variety of options to change the resulting report
-   Use YAML to set options for R Markdown
-   R Markdown supports a powerful citation system. Find more CSL styles at <https://github.com/citation-style-language/styles>
