---
title: "Visualizing Data"
output:
  html_document:
    df_print: paged
---

##  A quick introduction to Base R graphics

```{r}
mtcars
```

```{r}
hist(mtcars$mpg, breaks = 20)
```

```{r}
model <- lm(mpg ~ hp, data = mtcars)
  
plot(mtcars$hp, mtcars$mpg)
abline(model, lty = "dashed")
```

```{r}
par(mfrow = c(1, 2))     

plot(mtcars$mpg, mtcars$hwy,   
     col = mtcars$carb)  

plot(mtcars$mpg, mtcars$hwy,      
     col = mtcars$cyl)    
```



# Bonus

## Arranging Plots

You can arrange plots in many ways in R. Sometimes a facet does the job, but sometimes you need more control. ggplot2 uses the grid graphics system, and there are several packages that can manipulate grid objects. Two that are intended for ggplot2 are cowplot and patchwork. cowplot is good for when you need to make complicated changes to the plot but also includes `plot_grid()`, which creates a classic look often seen in journals. patchwork is a newer package (not yet on CRAN) that makes combining plots extremely simple: you just save the plots and add them together with `+`. Like cowplot, you can also make more complicated arrangements.

Let's try it with two plots.

```{r}
p1 <- ggplot(diabetes, aes(weight, hip, shape = gender)) +
  geom_jitter(size = 2.5) +
  theme_bw(base_size = 12) +
  theme(legend.position = "none") # remove legend from first plot

p2 <- ggplot(diabetes, aes(weight, hip, linetype = gender, shape = gender)) +
  geom_jitter(color = "grey90", size = 2.5) +
  geom_smooth(color = "black", se = FALSE) + 
  theme_bw(base_size = 12)
```

cowplot:

```{r}
library(cowplot)

plot_grid(p1, p2, labels = "auto")
```

patchwork:

```{r}
library(patchwork)

p1 + p2
```

## Spatial Graphics

R has many great libraries for working with spatial data. One that works particularly well is sf, the successor to the popular sp package. ggplot2 even has a geom for sf objects: `geom_sf()`. Let's look at SIDS data from North Carolina. This shape file is contained in the sf package. 

```{r}
library(sf)
nc <- st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

nc
```

Notice that it looks like a tibble with a lot of spatial information; that's because it is. It's designed to work well with dplyr and the other tidyverse tools.

```{r}
nc_plot <- nc %>% 
  gather("year", "rate", SID74, SID79) %>% 
  mutate(year = case_when(year == "SID74" ~ "1974",
                          year == "SID79" ~ "1979")) %>% 
  ggplot(aes(fill = rate)) +
    geom_sf(col = "white", size = .75) +
    scale_fill_viridis_c("Number of\nSIDS Deaths") +
    facet_wrap(~ year, ncol = 1) +
    theme_minimal(base_size = 16) 

nc_plot
```

Because coordinates are an essential aspect of ggplot2's design, it has great support for projections. 

```{r}
nc_plot + coord_sf(crs = st_crs(102003)) 
```

## Interactive Graphics

Likewise, R has a number of options for interactive graphics. These work in Rstudio but are best when included in an HTML RMarkdown document or a Shiny app. Popular packages for interactive visualizations are leaflet, mapview, highcharter, and plotly. plotly in particular makes it easy when you are using ggplot2. `ggplotly()` takes ggplot2 plots and automatically converts them to interactive graphs.

```{r}
library(plotly)
ggplotly(p2)
```

## Animated Graphics

gganimate is an in-development package for easy animation with ggplot2. Notice in the code below that only a couple layers in the ggplot2 code are new.

```{r}
library(gapminder)
library(gganimate)

p <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  #  facet_wrap(~continent) + # uncomment if you want to facet it by continent
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) + # from gganimate: what are the frames based on?
  ease_aes('linear') + # from gganimate: how should the transition flow?
  theme_minimal(base_size = 16) + 
  theme(legend.position = 'none')

animate(p, 100, 10)
```

