---
title: "<span style='color:#E69F00'>Data Visualization in R</span>"
subtitle: "<span style='color:#56B4E9'>ggplot2 fun & plotting in base R</span> <br>"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["default", "theme.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      navigation:
        scroll: false 
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, tibble.max_extra_cols = 6, tibble.width = 60)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center", dpi = 320, fig.height = 4)
library(tidyverse)
```

# Plots in Base R

```{r}
as_tibble(mtcars)
```

---

# `hist()`

```{r, fig.height=4}
hist(mtcars$mpg, breaks = 20)
```

---

# `plot()`

```{r, fig.height=4}
plot(mtcars$hp, mtcars$mpg)
```

---

# `plot()`

```{r, fig.height=4, eval=FALSE}
plot(mtcars$hp, mtcars$mpg)

model <- lm(mpg ~ hp, data = mtcars)
abline(model, lty = "dashed")
```

---

```{r, fig.height=4, echo=FALSE}
plot(mtcars$hp, mtcars$mpg)

model <- lm(mpg ~ hp, data = mtcars)
abline(model, lty = "dashed")
```

---

class: inverse, center

# Data Visualization Principles

---
class: inverse, center

# Data Visualization Principles

## **Simplify and focus**

---
class: inverse, center

# Data Visualization Principles

## ~~Simplify and focus~~
## **Increase your text size**

---
class: inverse, center

# Data Visualization Principles

## ~~Simplify and focus~~
## ~~Increase your text size~~
## **Avoid pie charts (use bar charts)**

---

class: inverse, center

# Data Visualization Principles

## ~~Simplify and focus~~
## ~~Increase your text size~~
## ~~Avoid pie charts (use bar charts)~~
## **Avoid line drawings (use fill/color)**

---

class: inverse, center

# Data Visualization Principles

## ~~Simplify and focus~~
## ~~Increase your text size~~
## ~~Avoid pie charts (use bar charts)~~
## ~~Avoid line drawings (use fill/color)~~
## **Stay 2D**

---
class: inverse

## **Bonus!**

--

1. Arranging plots

--

2. Spatial plots

--

3. Interactive plots

--

4. Animated plots

---

class: inverse, takeaway

## **`plot()`, `hist()`, and friends are useful for quick plots and sometimes working with packages**

## **ggplot2 is a powerful, extensible plotting system.**
