---
title: "<span style='color:#E69F00'>Data Visualization in R</span>"
subtitle: "<span style='color:#56B4E9'>ggplot2 fun & plotting in base R</span> <br>"
date: today
format: 
  kakashi-revealjs: 
    theme: [default, theme.css]
    highlight-style: github
    code-line-numbers: false
    scrollable: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, tibble.max_extra_cols = 6, tibble.width = 60)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center", dpi = 320, fig.height = 4)
library(tidyverse)
```

## Plots in Base R

```{r}
as_tibble(mtcars)
```

## `hist()`

```{r, fig.height=4}
hist(mtcars$mpg, breaks = 20)
```

## `plot()`

```{r, fig.height=4}
plot(mtcars$hp, mtcars$mpg)
```

## `plot()`

```{r, fig.height=4, eval=FALSE}
plot(mtcars$hp, mtcars$mpg)

model <- lm(mpg ~ hp, data = mtcars)
abline(model, lty = "dashed")
```

##

```{r, fig.height=4, echo=FALSE}
plot(mtcars$hp, mtcars$mpg)

model <- lm(mpg ~ hp, data = mtcars)
abline(model, lty = "dashed")
```

## Data Visualization Principles {.inverse .center}

## Data Visualization Principles {.inverse .center}

### **Simplify and focus**

## Data Visualization Principles {.inverse .center}

### ~~Simplify and focus~~
### **Increase your text size**

## Data Visualization Principles {.inverse .center}

### ~~Simplify and focus~~
### ~~Increase your text size~~
### **Avoid pie charts (use bar charts)**

## Data Visualization Principles {.inverse .center}

### ~~Simplify and focus~~
### ~~Increase your text size~~
### ~~Avoid pie charts (use bar charts)~~
### **Avoid line drawings (use fill/color)**

## Data Visualization Principles {.inverse .center}

### ~~Simplify and focus~~
### ~~Increase your text size~~
### ~~Avoid pie charts (use bar charts)~~
### ~~Avoid line drawings (use fill/color)~~
### **Stay 2D**

## **Bonus!** {.inverse}

::: {.incremental}
1. Arranging plots

2. Spatial plots

3. Interactive plots

4. Animated plots
:::

## {.inverse .takeaway}

### **`plot()`, `hist()`, and friends are useful for quick plots and sometimes working with packages**

### **ggplot2 is a powerful, extensible plotting system.**