---
title: "Reproduce Reports: figures"
format: html
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(ggdag, warn.conflicts = FALSE)
library(gapminder)
library(patchwork)
diabetes <- read_csv("diabetes.csv") |> 
  mutate(diabetic = ifelse(glyhb >= 6.5, "Diabetic", "Healthy"))
```

## Your Turn 1

1.  Using the chunk option defaults we discussed, set the global chunk options in the YAML metadata
2.  Render this document, and take a look at the first three figures. Do you like how they look?
3.  Modify `fig.width` for each chunk until you're satisfied

```{r}
#| label: fig-gapminder
#| fig.cap: A lineplot of life expectancy over time by country
ggplot(gapminder, aes(year, lifeExp, color = country)) + 
  geom_line() + 
  scale_color_manual(values = country_colors) +
  theme_minimal(14) +
  theme(legend.position = "none") +
  labs(
    x = "",
    y = "Life expectancy"
  )
```

```{r}
#| label: fig-ggdag
#| fig.cap: A causal diagram illustrating butterfly bias
ggdag(butterfly_bias()) + 
  theme_dag()
```

```{r}
#| label: fig-diabetes
#| fig.cap: Understanding the relationship between diabetes and hip/waist ratio
fig1a <- diabetes |> 
  drop_na() |> 
  ggplot(aes(ratio, glyhb)) + 
  geom_point(shape = 21, fill = "grey80", color = "white", size = 2) + 
  geom_smooth(
    size = 1, 
    color = "steelblue", 
    se = FALSE, 
    method = "lm", 
    formula = y ~ x
  ) + 
  scale_x_log10() +
  theme_minimal(14) +
  labs(
    x = "Hip/waist ratio",
    y = "Hemoglobin A1c",
    tag = "A"
  )

fig1b <- diabetes |> 
  drop_na() |> 
  mutate(diabetic = factor(diabetic, levels = c("Healthy", "Diabetic"))) |> 
ggplot(aes(ratio, fill = diabetic)) + 
  geom_density(color = "white", alpha = .8) +
  theme_minimal(14) +
  theme(
    legend.title = element_blank(),
    legend.position = "bottom"
  ) +
  labs(x = "Hip/waist ratio", tag = "B") +
  scale_fill_manual(values = c("steelblue", "firebrick"))

fig1a + fig1b
```

## Your Turn 2

1.  Globally set `fig.path` to "figures/figure-". This will tell knitr to create figures in the "figures" folder with a prefix of "figures-".
2.  Render this document and take a look at the images in the `figures` folder.

## Your Turn 3

1.  Cross-reference one of the figures above.

<!-- Describe a figure below! -->

<!-- These are HTML comments, by the way. Quarto ignores them! -->

## Your Turn 4

1.  Include `external_img/r_rollercoaster.png` in the code chunk below.
2.  Render
3.  Let's change a few chunk options:

-   Add a chunk label
-   Set `fig.alt` describing the image
-   Modify `out.width` to use a different percentage than the default.

4.  Render again

```{r}

```

------------------------------------------------------------------------

# Take aways

-   Getting figures to output exactly as you want is iterative, but a few sensible defaults can help
-   When you use the suggested defaults, leverage `fig.width` to scale text and line size
-   `ggsave()` and `ragg::agg_png()` also have scaling options
-   ragg helps make crisper, more consistent plots across operating systems. Set it as your default in knitr with the chunk option `dev = "ragg_png"` and in the RStudio Viewer via `Tools > Global Options > Graphics > Backend > AGG`
-   Use `knitr::include_graphics()` to include external graphics
