---
title: "Tidy Tuesday"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# knit options
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Exercise

The purpose of this exercise is to practice your R wrangling and visualization skills with real data sets from the Tidy Tuesday project:

> The intent of Tidy Tuesday is to provide a safe and supportive forum for individuals to practice their wrangling and data visualization skills independent of drawing conclusions. While we understand that the two are related, the focus of this practice is purely on building skills with real-world data.

The Tidy Tuesday project shares data every week for people to cultivate data science skills. Some people post on Twitter, on the #TidyTuesday hashtag, some share on GitHub, and some just work privately. It's a great resource for data when you're looking to practice! 

In this project, we'll start with a guided exploration of a TT dataset. Then, you'll pick a dataset to practice exploratory data analysis on.

## Part 1: European Energy

We're going to explore a dataset about energy use in Europe from 2016 to 2018.  This was the Tidy Tuesday dataset the week of 2020-08-04. You can find more details here: https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-08-04/readme.md

First, let's read in the data.

```{r}
energy_types <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-08-04/energy_types.csv")
country_totals <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-08-04/country_totals.csv")
```

```{r}
energy_types <- energy_types %>%
  filter(level == "Level 1") %>% 
  mutate(country_name = ifelse(is.na(country_name), "UK", country_name))
```

```{r}
library(cowplot)

energy_types_longer <- energy_types %>%
  pivot_longer(
    cols = `2016`:`2018`, 
    names_to = "year", 
    values_to = "gigawatts"
  ) 

energy_types_longer %>% 
  ggplot(aes(x = year, y = gigawatts, group = country)) +
  geom_line(color = "steelblue", size = 0.8) + 
  geom_point(shape = 21, fill = "steelblue", color = "white", size = 2) +
  facet_wrap(vars(type), scales = "free_y") +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal_hgrid()
```

```{r}
wind_2018 <- energy_types_longer %>% 
  filter(year == "2018", type == "Wind") %>% 
  arrange(gigawatts) %>% 
  mutate(country_name = fct_inorder(country_name))

wind_2018 %>% 
  ggplot(aes(x = gigawatts, y = country_name)) +
  geom_col(fill = "steelblue", alpha = .9) +
  scale_x_continuous(labels = scales::comma) +
  ylab(NULL) +
  theme_minimal_vgrid()
```


```{r}
library(ggtext)

wind_2018 %>% 
  mutate(
    germany = country_name == "Germany",
    country_name = fct_recode(country_name, "<span style='color:firebrick'>Germany</i>" = "Germany")
  ) %>% 
  ggplot(aes(x = gigawatts, y = country_name)) +
  geom_col(aes(fill = germany), alpha = .9, show.legend = FALSE) +
  scale_fill_manual(values = c("steelblue", "firebrick")) + 
  scale_x_continuous(labels = scales::comma) +
  ylab(NULL) +
  theme_minimal_vgrid() +
  theme(axis.text.y = element_markdown())
```

```{r}
energy_types %>% 
  filter(type == "Conventional thermal") %>% 
  mutate(change = `2018` - `2016`) %>% 
  arrange(desc(change)) %>% 
  mutate(country_name = fct_inorder(country_name)) %>% 
  ggplot(aes(x = 0, xend = change, y = country_name, yend = country_name, color = change > 0)) +
  geom_segment(arrow = grid::arrow(length = unit(1, "mm"), type = "closed")) +
  labs(
    y = NULL,
    x = "Change in conventional thermal\nproduction from 2016 to 2018 (gigawatts)"
  ) +
  scale_color_manual(values = c("firebrick", "steelblue")) + 
  theme_minimal_vgrid() +
  theme(legend.position = "none")
```


## Part 2: Analyze a Tidy Tuesday dataset 

In the second part of this assignment, I want you to pick a different Tidy Tuesday dataset. Find a dataset you are interested in here: https://github.com/rfordatascience/tidytuesday#datasets. Then, click on the link in the "data" column to go to the README file for the dataset, which will tell you how to import the data. Generally, the instructions will offer several options for importing; any of them are fine. In part 1, I just used readr and a link to repository.

How you explore your data is up to you! Remember the whole toolkit you have at your disposal: wrangling, visualizing, tidying, joins, and so on. You'll likely need to use a combination of some or all of these tools for any given question.

### Questions

1. Describe the data set in your own words (1 paragraph). 
2. Produce 2-3 graphs in ggplot2 that show something interesting in the data. You'll have to to some exploratory analysis to find something, so make sure to include that code. The plots can be about anything using any type of geoms you like. 

*Requirements*:

* The graph should attempt to answer a specific question you have about the data. Clearly state the question the graph is trying to answer, perhaps as a title or in a short description of the plot.
* All labels should be clear and easy to read. That may mean that you need to replace, for example, the axes labels with descriptions rather than the name of the variable in the data set (e.g. I don't want to see "some_weird_variable_name2")
* Likewise, all content on the graph should be reasonably legible
* Include both the code (in this R Markdown file or in a separate R script) and a `.png` file created with `ggsave()`. 

Note: Tidy Tuesday is a community event, so it's okay to draw inspiration from the work of others. *However, you must include a link to your inspiration, and you can't just copy and paste someone else's code.* You can find other people's work for Tidy Tuesday on the #TidyTuesday hashtag on Twitter. Another useful resources is David Robinson's Tidy Tuesday livestream, where he analyzes data live without previously looking at the data for the week: https://www.youtube.com/channel/UCeiiqmVK07qhY-wvg3IZiZQ
