---
title: "R Markdown: Whole Game"
output: html_document
---

For this set of exercises, we'll work through a targets pipeline to run an analysis and generate an R Markdown report in `whole_game.Rmd`.
Open `_targets.R`---our file to manage the targets pipeline---to get started.

```{r setup}
library(targets)

copy_targets_template <- function(template) {
  copied <- file.copy(
    file.path("templates", template),
    "_targets.R",
    overwrite = TRUE
  )
  
  if (!copied) {
    stop(
      file.path("templates", template), " not found. ", 
      "Are you sure you're in the right working directory?",
      call. = FALSE
    )
  }
  
  template_r <- usethis::ui_path(file.path("templates", template))
  targets_r <- usethis::ui_path("_targets.R")
  usethis::ui_done("Copying {template_r} to {targets_r}")
}
```


## Your Turn 1

Run this code to make a fresh copy of `_targets.R`:

```{r}
# this function is defined in the `setup` chunk above!
copy_targets_template("your_turn_1.R")
```

1.  Open `_targets.R` and take a look at the structure of the file. targets requires this file to run! You can also open this file at the console with `tar_edit()`.
2.  Add "tidyverse" to the `packages` argument of `tar_option_set()`
3.  We're going to add two targets to `tar_plan()`. First, add `tar_file(forest_file, "data/forest.csv")`. This creates a new target called `forest_file`: `tar_file()` essentially tells targets that "data/forest.csv" is a file we should keep an eye on.
4.  Next, add `forest = read_csv(forest_file)`
5.  Then, run `tar_visnetwork()` in the console (don't include this in your script). What do you see?
6.  Run `tar_make()` in the console (not in your script)
7.  In the console, run `tar_read(forest)`. What are you seeing?

## Your Turn 2

Get a little lost on the last exercise? Run this code to update `_targets.R` and catch up:

```{r}
copy_targets_template("your_turn_2.R")
```

1.  Open `R/functions.R` and take a look at the first function, `map_2010_forest_conversion()`. This function takes a dataframe---forest---and generates a map of forest conversion in 2010.
2.  Let's add this map to our targets plan. First, add a few more packages that we need for this function to the `packages` argument of `tar_option_set()`: "maps", "sf", "scico".
3.  We also need to load our custom functions. Add `source("R/functions.R")` after `tar_option_set()`.
3.  Add `forest_map = map_2010_forest_conversion(forest)` as a target
4.  Run `tar_visnetwork()` in the console. What's changed? Why are you seeing other functions beyond what you added to the plan?
5.  Run `tar_make()` in the console.
6.  Run `tar_read(forest_map)` in the console.

## Your Turn 3

Get a little lost on the last exercise? Run this code to update `_targets.R` and catch up:

```{r}
copy_targets_template("your_turn_3.R")
```

1.  The next function in `R/functions.R` , `barplot_top_10_change()`, takes `forest` and returns a barplot of the 10 countries with the highest absolute change in 2010. Add this function as a target called `top10_barplot`.
2.  Take a look with `tar_visnetwork()`, then run `tar_make()`.
3.  Use `tar_read()` to look at the the bar plot

## Your Turn 4

Get a little lost on the last exercise? Run this code to update `_targets.R` and catch up:

```{r}
copy_targets_template("your_turn_4.R")
```

1.  Add patchwork to the packages argument at the top of `_targets.R`
2.  Remove the `top10_barplot` target and change the `forest_map` target to use `plot_top_change_2010(forest)` instead.
3.  Run `tar_visnetwork()`, `tar_make()`, and `tar_read()`

## Your Turn 5

Get a little lost on the last exercise? Run this code to update `_targets.R` and catch up:

```{r}
copy_targets_template("your_turn_5.R")
```

1.  One at a time, running `tar_visnetwork()` , `tar_make()`, and `tar_read()` at each step, add five new targets:

    -   Add "data/brazil_loss.csv" as a file target called `brazil_loss_file`

    -   Read `brazil_loss_file` using `read_csv()`.
        Call the target `brazil_loss`

    -   Add a new target called `brazil_causes_2010` using `clean_2010_causes()` on `brazil_loss`

    -   Add a new target called `brazil_causes_table` using `table_brazil_causes()` on `brazil_causes_2010`. For this target, we also need to add "gt" to the packages in `tar_option_set()` since we use the gt package for this table.

    -   Add a new target called `brazil_causes_hist` using `plot_cause_hist()` on `brazil_loss`. Note that when you look at ggplots, ggplot2 actually re-builds the plot on the fly. That means you may need to load the packages and functions we're using in your local environment, as well.

## Your Turn 6

Get a little lost on the last exercise? Run this code to update `_targets.R` and catch up:

```{r}
copy_targets_template("your_turn_6.R")
```

1.  Open `whole_game.Rmd` and take a look. This is a report that uses the targets we created in our pipeline.
2.  Use `tar_load()` to load `forest` and `brazil_causes_2010` in the `setup` chunk. This time, you will include `tar_load()` directly in your code!
3.  Use `tar_read()` in the following chunks: `forest-map`, `causes-table`, and `causes-hist`.
4.  Add the report to your pipeline in `_targets.R`. The format for for rendering a R Markdown file is `tar_render(name_of_target, "name_of_report.Rmd")`.
5.  Run `tar_visnetwork()` and `tar_make()`. Open your rendered report from the Files pane.

## Your Turn 7

```{r}
copy_targets_template("your_turn_7.R")
```

1.  Run `tar_destroy()` in the console
2.  Run `tar_visnetwork()`. What's different?
3.  Run `tar_make()`
4.  Take a look at your rendered report
