---
title: "Whole Game: Analyzing deforestation"
format: html
bibliography: packages.bib
nocite: |
  @*
---

## Introduction

For this module, we'll focus on generating a report using targets.
We'll create each result as a series of targets, including them in `_targets.R`.
We'll iterate through, one or two targets at a time, running `tar_make()` and `tar_visnetwork()` to check our work.
Check `exercises.qmd` for guidance on creating this report!
This will be an intentionally simplified report: we'll focus on the basics of creating output as targets and including them here as rendered material.

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = FALSE, 
  dev = "ragg_png",
  dpi = 320, 
  out.width = "80%",
  fig.width = 6,
  fig.asp = 0.618,
  fig.retina = 2,
  fig.align = "center",
  fig.show = "hold"
)

library(tidyverse)
library(targets)
source("R/functions.R")

# write a bibliography file for all the packages we're using
knitr::write_bib(file = "packages.bib")

# fill me in
________(forest)
________(brazil_causes_2010)
```

## Methods

In this analysis, we look at the net change in forestation in 2010 as well as the causes of deforestation in Brazil.
These data come from [Our World in Data](https://ourworldindata.org/forests-and-deforestation).

## Results

```{r}
#| label: top-forest-changes
top_changes <- forest |> 
  filter(year == 2010, entity != "World") |> 
  slice_max(abs(net_forest_conversion), n = 2)

pull_net <- function(top_changes, country) {
  top_changes |> 
    filter(entity == country) |> 
    pull(net_forest_conversion) |> 
    scales::comma()
}
```

```{r}
#| label: fig-forest-map
#| fig-cap: "Net Forest Conversion in 2010"
#| fig-width: 9
________(forest_map)
```

@fig-forest-map presents a map of net forest conversion in 2010, as well as the 10 countries with the highest absolute change in forest.
In 2010, the China gained more forest than any other country, a net growth of `r pull_net(top_changes, "China")` hectacres; Brazil lost more forest than any other country, a net loss of `r pull_net(top_changes, "Brazil")` hectacres.

```{r}
#| label: tbl-causes-table
#| tbl-cap: "The causes of deforestation in Brazil in 2010."
pull_hect <- function(brazil_causes_2010, reason) {
  brazil_causes_2010 |> 
    filter(cause == reason) |> 
    pull(hectacres) |> 
    scales::comma()
}

________(brazil_causes_table)
```

@tbl-causes-table shows the causes of deforestation in Brazil in 2010.
The top cause was clearing for pasture (`r pull_hect(brazil_causes_2010, "Pasture")` hectacres), followed by fire (`r pull_hect(brazil_causes_2010, "Fire")` hectacres), then small scale clearing (`r pull_hect(brazil_causes_2010, "Small scale clearing")` hectacres).

```{r}
#| label: fig-causes-hist
#| fig-cap: "Causes of deforestation in Brazil over time, sorted by maximum loss"
#| fig-width: 9
________(brazil_causes_hist)
```

@fig-causes-hist shows that, while deforestation has decreased in Brazil since the early 2000s, it remains steady at over 1 million hectacres lost per year.
Pasture clearing is consistently the top cause, while fire---generally being accidental---varies from year to year.

## Discussion

We've combined the magic of Quarto for generating reports with the reliability of targets for managing our code, when we render the report, and understanding what targets the report depends on.
targets also let's us focus more on communicating and less on including lots of code in our report; it's all handled by our pipeline.
Using `tar_load()` and `tar_read()` allows us to grab targets and associated them with this report!

## References
