---
title: "Types and Functions"
output: html_document
---

```{r setup}
library(tidyverse)
library(gapminder)
```

## Your Turn 1

1. Create a character vector of colors using `c()`. Use the colors "grey90" and "steelblue". Assign the vector to a name. 2. Use the vector you just created to change the colors in the plot below using `scale_color_manual()`. Pass it using the `values` argument.

```{r}
cols <- c("grey90", "steelblue")

gapminder %>% 
  mutate(rwanda = ifelse(country == "Rwanda", TRUE, FALSE)) %>% 
  ggplot(aes(year, lifeExp, color = rwanda, group = country)) + 
    geom_line() +
    scale_color_manual(values = cols) +
    theme_minimal()
```

## Your Turn 2

1. Create a numeric vector that has the following values: 3, 5, NA, 2, and NA. 
2. Try using `sum()`. Then add `na.rm = TRUE`.
3. Check which values are missing with `is.na()`; save the results to a new object and take a look
4. Chance all missing values of `x` to 0
5. Try `sum()` again without `na.rm = TRUE`.

```{r}
x <- c(3, 5, NA, 2, NA)
sum(x)
sum(x, na.rm = TRUE)

x_missing <- is.na(x)
x_missing

x[x_missing] <- 0
sum(x)
```

## Your Turn 3

1. Create a function called `sim_data` that doesn't take any arguments.
2. In the function body, we'll return a `tibble`. 
3. For `x`, have `rnorm()` return 50 random numbers.  
4. For `sex`, use `rep()` to create 50 values of "male" and "female". Hint: You'll have to give `rep()` a character vector. for the first argument. The `times` argument is how many times `rep()` should repeat the first argument, so make sure you 3. account for that.
5. For `age()` use the `sample()` function to sample 50 numbers from 25 to 50 with replacement.
6. Call `sim_data()`

```{r}
sim_data <- function() {
  tibble(
    x = rnorm(50),
    sex = rep(c("male", "female"), times = 25),
    age = sample(25:50, size = 50, replace = TRUE)
  )
}

sim_data()
```

## Your Turn 3: Bonus!

Did you finish early? Try this extra exercise

1. Set a seed for the random number generator using `set.seed()`. Check the help page if you're not sure how.
2. Call `sim_data()`
3. Press the green play button several times. What's happening here?

```{r}

```

## Your Turn 4

1. Write a function to calculate an E-Value given an RR. Call the argument `estimate`. In the body of the function, calculate the E-Value using `estimate + sqrt(estimate * (estimate - 1))`
2. Call `evalue()` for a risk ratio of 3.9

```{r}
evalue <- function(estimate) {
  estimate + sqrt(estimate * (estimate - 1))
}

evalue(3.9)
```

# Your Turn 5

1. Use `if ()` together with `is.numeric()` to make sure `estimate` is a number. Remember to use `!` for not.
2. If the estimate is less than 1, set `estimate` to be equal to `1 / estimate`.
3. Call `evalue()` for a risk ratio of 3.9. Then try 0.80

```{r}
evalue <- function(estimate) { 
  if (!is.numeric(estimate)) stop("`estimate` must be numeric")
  if (estimate < 1) estimate <- 1 / estimate
  estimate + sqrt(estimate * (estimate - 1))
}

evalue(3.9)
evalue(.80)
evalue("3.9")
```

# Your Turn 6

1. Add a new argument called `type`. Set the default value to "rr"
2. Check if `type` is equal to "or". If it is, set the value of `estimate` to be `sqrt(estimate)`
3. Call `evalue()` for a risk ratio of 3.9. Then try it again with `type = "or"`.

```{r}
evalue <- function(estimate, type = "rr") {
  if (!is.numeric(estimate)) stop("`estimate` must be numeric")
  if (type == "or") estimate <- sqrt(estimate)
  if (estimate < 1) estimate <- 1 / estimate
  estimate + sqrt(estimate * (estimate - 1))
}

evalue(3.9)
evalue(3.9, type = "or")
```



# Your Turn 7: Challenge!

1. Create a new function called `transform_to_rr` with arguments `estimate` and `type`.
2. Use the same code above to check if `type == "or"`. Add another line that checks if `type == "hr"`. If it does, transform the estimate using this formula: `(1 - 0.5^sqrt(estimate)) / (1 - 0.5^sqrt(1 / estimate))`.
3. Move the code that checks if `estimate < 1` below the OR and HR transformations
4. Return `estimate`
5. In `evalue()`, change the default argument of `type` to be a character vector containing "rr", "or", and "hr".
6. Get and validate the value of `type` using `match.arg()`. Follow the pattern `argument_name <- match.arg(argument_name)`
7. Transform `estimate` using `transform_to_rr()`. Don't forget to pass it both `estimate` and `type`!

```{r}
transform_to_rr <- function(estimate, type) {
  if (type == "or") estimate <- sqrt(estimate)
  if (type == "hr") {
    estimate <- (1 - 0.5^sqrt(estimate)) / (1 - 0.5^sqrt(1 / estimate))
  }
  if (estimate < 1) estimate <- 1 / estimate
    
  estimate
}

evalue <- function(estimate, type = c("rr", "or", "hr")) {
  # validate arguments
  if (!is.numeric(estimate)) stop("`estimate` must be numeric")
  type <- match.arg(type)

  # calculate evalue
  estimate <- transform_to_rr(estimate, type)
  estimate + sqrt(estimate * (estimate - 1))
}

evalue(3.9)
evalue(3.9, type = "or")
evalue(3.9, type = "hr")
evalue(3.9, type = "rd")
```


## Your Turn 8

1. Use `...` to pass the arguments of your function, `filter_summarize()`, to `filter()`.
2. In summarize, get the n and mean life expectancy for the data set
3. Check `filter_summarize()` with `year == 1952`. 
4. Try `filter_summarize()` again for 2002, but also filter countries that have "and" in the country name. Use `str_detect()` from the stringr package.

```{r}
filter_summarize <- function(...) {
  gapminder %>% 
    filter(...) %>% 
    summarize(n = n(), mean_lifeExp = mean(lifeExp))
}

filter_summarize(year == 1952)
filter_summarize(year == 2002, str_detect(country, " and "))
```


## Your turn 9

1. Filter gapminder by `.year`.
2. Arrange it by the variable. Don't forget to wrap it in curly-curly!
3. Make a scatter plot. Use `variable` for x. For y, we'll use `country`, but to keep it in the order we arranged it by, we'll turn it into a factor. Wrap the the `factor()` call with `fct_inorder()`. Check the help page if you want to know more about what this is doing.

```{r}
top_barplot <- function(variable, .year) {
  gapminder %>%
    filter(year == .year) %>% 
    arrange(desc({{variable}})) %>% 
    #  take the 10 lowest values
    tail(10) %>% 
    ggplot(aes(x = {{variable}}, y = fct_inorder(factor(country)))) +
      geom_point() +
      theme_minimal()
}

top_barplot(lifeExp, 2002)
top_barplot(lifeExp, 2002) + labs(x = "Country", y = "Life Expectancy")
```

## Your turn 9: Bonus! 

Finish early? Try this exercise.

1. Filter by `year` by the value of `.year`
2. Group by `continent`
3. Summarize the number of countries per continent. Call the new variable `n_countries`
4. Arrange the data `n_countries`
5. Make a bar plot with `y` set to `n_countries` and `x` set to continent. Make `continent` an in-order factor using the pattern above (`fct_inorder(factor(var))`). Use `geom_col()` instead of `geom_bar()` since you are specifying two variables.
6. Set the theme to `theme_minimal()`. 
7. Set the axis title for y to `element_blank()` using `theme()`
8. Use `paste()` to make a label for the y axis. (Wait, didn't we just turn that off? Hang on...)
9. Rotate the plot with `coord_flip()`
10. Finally, plot the number of countries per continent for 2002

```{r}
plot_n_countries <- function(.year) {
  gapminder %>% 
    filter(year == .year) %>% 
    group_by(continent) %>% 
    summarise(n_countries = n()) %>% 
    arrange(n_countries) %>% 
    ggplot(aes(y = n_countries, x = fct_inorder(factor(continent)))) + 
      geom_col() +
      theme_minimal() +
      theme(axis.title.y = element_blank()) +
      ylab(paste("n Countries in", .year)) + 
      coord_flip()
}

plot_n_countries(2002)
```



***

# Take aways

* Vectors in R each have a type. These include numeric, integer, character, factor, logical, and date. 
* You can subset vectors or modify them using brackets like `x[[1]]`
* Writing functions in R follows this pattern:

```
function_name <- function(arg1, arg2 = NULL) {
    # compute something to return
}
```

* Functions are objects in R like everything else!
* To program with tidyverse functions, use `...` or curly-curly in the pattern of `{{x}}`. This lets us use bare names like we normally do with these functions.
