---
title: "Visualizing Data"
output:
  html_document:
    df_print: paged
---

```{r setup}
library(tidyverse)
diabetes <- read_csv("diabetes.csv")
```

## Your Turn 10

1. Change the color scale by adding a scale layer. Experiment with `scale_color_distiller()` and `scale_color_viridis_c()`. Check the help pages for different palette options.
2. Set the color aesthetic to `gender`. Try `scale_color_brewer()`.
3. Set the colors manually with `scale_color_manual()`. Use `values = c("#E69F00", "#56B4E9")` in the function call.
4. Change the legend title for the color legend. Use the `name` argument in whatever scale function you're using.

```{r}
diabetes %>% 
  ggplot(aes(waist, hip, col = weight)) + 
  geom_point() +
  scale_color_distiller()

diabetes %>% 
  ggplot(aes(waist, hip, col = gender)) + 
  geom_point() +
  scale_color_brewer()

diabetes %>% 
  ggplot(aes(waist, hip, col = gender)) + 
  geom_point() +
  scale_color_manual(values = c("#E69F00", "#56B4E9"))


diabetes %>% 
  ggplot(aes(waist, hip, col = gender)) + 
  geom_point() +
  scale_color_manual(name = "Sex", values = c("#E69F00", "#56B4E9"))
```

## Your Turn 11

1. Change the theme using one of the built-in theme functions. 
2. Use `theme()` to change the legend to the bottom with `legend.position = "bottom"`.
3. Remove the axis ticks by setting the `axis.ticks` argument to `element_blank()`
4. Change the font size for the axis titles. Use `element_text()`. Check the help page if you don't know what option to change.

```{r}
diabetes %>% 
  ggplot(aes(waist, hip, col = weight)) + 
  geom_point() +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.ticks = element_blank(),
    axis.title = element_text(size = 16)
  )
```

## Your Turn 12

Take a look at the `diamonds` data set from ggplot2. How many rows does it have?
```{r}
diamonds
```

1. Make a scatterplot of `carat` vs. `price`. How's it look?
2. Try adjusting the transparency. 
3. Replace `geom_point()` with 2d bins.
4. Try hex bins.

```{r}
diamonds %>% 
  ggplot(aes(carat, price)) + 
    geom_point(alpha = .2)

diamonds %>% 
  ggplot(aes(carat, price)) + 
    geom_bin2d()

diamonds %>% 
  ggplot(aes(carat, price)) + 
    geom_hex()
```

## Your Turn 13

1. Add a title. 
2. Change the x and y axis labels to include the unites (inches for `hip` and pounds for `weight`). You can use either `labs()` or `xlab()` and `ylab()`
3. Add `scale_linetype()` and set the `name` argument to "Sex".

```{r}
ggplot(diabetes, aes(weight, hip, linetype = gender)) +
  geom_jitter(alpha = .2, size = 2.5) +
  geom_smooth(color = "black", se = FALSE) + 
  theme_bw(base_size = 12) +
  ggtitle("My title") +
  xlab("Weight (inches)") + 
  ylab("Hip (lbs)") +
  scale_linetype(name = "Sex")
  
```


## Your Turn 14

Save the last plot and then locate it in the files pane.

```{r}
ggsave(filename = "plot.png")
```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**. You may not even need to add scales and themes!

```{r eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>)) +
  <SCALE_FUNCTION>() + 
  <THEME_FUNCTION>()
```


# Bonus

## Arranging Plots

You can arrange plots in many ways in R. Sometimes a facet does the job, but sometimes you need more control. ggplot2 uses the grid graphics system, and there are several packages that can manipulate grid objects. Two that are intended for ggplot2 are cowplot and patchwork. cowplot is good for when you need to make complicated changes to the plot but also includes `plot_grid()`, which creates a classic look often seen in journals. patchwork is a newer package (not yet on CRAN) that makes combining plots extremely simple: you just save the plots and add them together with `+`. Like cowplot, you can also make more complicated arrangements.

Let's try it with two plots.

```{r}
p1 <- ggplot(diabetes, aes(weight, hip, shape = gender)) +
  geom_jitter(size = 2.5) +
  theme_bw(base_size = 12) +
  theme(legend.position = "none") # remove legend from first plot

p2 <- ggplot(diabetes, aes(weight, hip, linetype = gender, shape = gender)) +
  geom_jitter(color = "grey90", size = 2.5) +
  geom_smooth(color = "black", se = FALSE) + 
  theme_bw(base_size = 12)
```

cowplot:

```{r}
library(cowplot)

plot_grid(p1, p2, labels = "auto")
```

patchwork:

```{r}
# patchwork is not on CRAN yet
# install it with:
# remotes::install_github("thomasp85/patchwork")
library(patchwork)

p1 + p2
```

## Spatial Graphics

R has many great libraries for working with spatial data. One that works particularly well is sf, the successor to the popular sp package. ggplot2 even has a geom for sf objects: `geom_sf()`. Let's look at SIDS data from North Carolina. This shape file is contained in the sf package. 

```{r}
library(sf)
nc <- st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

nc
```

Notice that it looks like a tibble with a lot of spatial information; that's because it is. It's designed to work well with dplyr and the other tidyverse tools.

```{r}
nc_plot <- nc %>% 
  gather("year", "rate", SID74, SID79) %>% 
  mutate(year = case_when(year == "SID74" ~ "1974",
                          year == "SID79" ~ "1979")) %>% 
  ggplot(aes(fill = rate)) +
    geom_sf(col = "white", size = .75) +
    scale_fill_viridis_c("Number of\nSIDS Deaths") +
    facet_wrap(vars(year), ncol = 1) +
    theme_minimal(base_size = 16) 

nc_plot
```

Because coordinates are an essential aspect of ggplot2's design, it has great support for projections. 

```{r}
nc_plot + coord_sf(crs = st_crs(102003)) 
```

## Interactive Graphics

Likewise, R has a number of options for interactive graphics. These work in Rstudio but are best when included in an HTML RMarkdown document or a Shiny app. Popular packages for interactive visualizations are leaflet, mapview, highcharter, and plotly. plotly in particular makes it easy when you are using ggplot2. `ggplotly()` takes ggplot2 plots and automatically converts them to interactive graphs.

```{r}
library(plotly)
ggplotly(p2)
```

## Animated Graphics

gganimate is an in-development package for easy animation with ggplot2. Notice in the code below that only a couple layers in the ggplot2 code are new.

```{r}
library(gapminder)
library(gganimate)

p <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  #  facet_wrap(vars(continent)) + # uncomment if you want to facet it by continent
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) + # from gganimate: what are the frames based on?
  ease_aes('linear') + # from gganimate: how should the transition flow?
  theme_minimal(base_size = 16) + 
  theme(legend.position = 'none')

animate(p, 100, 10)
```

