---
title: "Visualizing Data"
output:
  html_document:
    df_print: paged
---

```{r setup}
library(tidyverse)
```

## Your Turn 1

Read in the `diabetes` data.

```{r}
diabetes <- read_csv("diabetes.csv")
```

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}
ggplot(data = diabetes, mapping = aes(x = weight, y = hip)) +
  geom_point()
```

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph using the `gender` variable. Experiment.  

```{r}
ggplot(data = diabetes, mapping = aes(x = weight, y = hip, color = gender)) +
  geom_point()
ggplot(data = diabetes, mapping = aes(x = weight, y = hip, size = gender)) +
  geom_point()
ggplot(data = diabetes, mapping = aes(x = weight, y = hip, alpha = gender)) +
  geom_point()
ggplot(data = diabetes, mapping = aes(x = weight, y = hip, shape = gender)) +
  geom_point()

ggplot(data = diabetes, mapping = aes(x = weight, y = hip)) +
  geom_point(aes(color = gender, shape = gender))
```

Try moving the mapping argument to `geom_point()`. Add in any aesthetics you found helpful.

```{r}
ggplot(data = diabetes, mapping = aes(x = weight, y = hip)) +
  geom_point()
```

## Your Turn 3

Replace this scatterplot with one that draws boxplots. Try typing "geom_" and hitting the tab button to see the functions that start with "geom_". Try to guess which one to use. If you need help, there's a copy of the cheatsheet in this folder.

```{r}
ggplot(diabetes, aes(gender, chol)) + geom_boxplot()
```

## Your Turn 4

Make a histogram of the `glyhb` variable in `diabetes`.

```{r}
ggplot(diabetes, aes(x = glyhb)) + geom_histogram()
```

Redo the `glyhb` plot as a density plot.

```{r}
ggplot(diabetes, aes(x = glyhb)) + geom_density()

```

---

## Your Turn 5

Make a bar chart `frame` colored by `gender`. Then, try it with the `fill` aesthetic instead of `color`.

```{r}
diabetes %>% 
  drop_na() %>% 
  ggplot(aes(frame, color = gender)) + 
    geom_bar()

diabetes %>% 
  drop_na() %>% 
  ggplot(aes(frame, fill = gender)) + 
    geom_bar()
```

## Your Turn 6

Take your code for the bar chart above (using the `fill` aesthetic). Experiment with different `position` values: "dodge", "fill", "stack"

```{r}
diabetes %>% 
  drop_na() %>% 
  ggplot(aes(frame, fill = gender)) + 
    geom_bar(position = "dodge")

diabetes %>% 
  drop_na() %>% 
  ggplot(aes(frame, fill = gender)) + 
    geom_bar(position = "fill")

diabetes %>% 
  drop_na() %>% 
  ggplot(aes(frame, fill = gender)) + 
    geom_bar(position = "stack")
```

## Your Turn 7

Run the code after every change you make.

1. Predict what this code will do. Then run it. 
2. Add a `linetype` aesthetic for `gender`. Run it again.
3. *Set* the color of `geom_smooth()` to "black"
4. Add `se = FALSE` to the `geom_smooth()`
5. It's hard to see the lines well now. How about setting `alpha = .2` in `geom_point()`?
6. Jitter the points. You can either change the geom or change the `position` argument.
7. Add another layer, `theme_bw()`. Remember to use `+`.

```{r}
ggplot(diabetes, aes(weight, hip, linetype = gender)) + 
  geom_point(alpha = .2, position = "jitter") +
  geom_smooth(color = "black", se = FALSE) +
  theme_bw()
```

## Your Turn 8

Use a facet grid by `gender` and `location`

```{r}
ggplot(diabetes, aes(weight, hip)) + 
  geom_point() +
  geom_smooth() +
  facet_grid(vars(gender), vars(location))
```

## Your Turn 9: Challenge!

1. Load the `datasauRus` package. This package includes a data set called `datasaurus_dozen`.
2. Use dplyr to summarize the correlation between `x` and `y`. First, group it by `dataset`, and then summarize with the `cor()` function. Call the new variable `corr`. What's it look like?
3. Mutate `corr`. Round it to 2 digits. Then, mutate it again (or wrap it around your first change) using: `paste("corr:", corr)`
4. Save the summary data frame as `corrs`.

```{r}
library(datasauRus)
corrs <- datasaurus_dozen %>% 
  group_by(dataset) %>% 
  summarize(
    corr = cor(x, y)
  ) %>% 
  mutate(
    corr = round(corr, 2),
    corr = paste("corr:", corr)
  )

corrs
```

5. Pass `datasaurus_dozen` to `ggplot()` and add a point geom
6. Use a facet (wrap) for `dataset`.
7. Add a text geom. For this geom, set `data = corrs`. You also need to use `aes()` in this call to set `label = corr`, `x = 50`, and `y = 110`.

```{r}
datasaurus_dozen %>% 
  ggplot(aes(x, y)) +
    geom_point() + 
    facet_wrap(vars(dataset)) +
    geom_text(data = corrs, aes(label = corr, x = 50, y = 110))

```

