---
title: "ggplot2: Plot replication challenge!"
output:
  html_document:
    df_print: paged
---

```{r setup}
# Load your packages!
library(tidyverse)
library(ggrepel)
```

# Exercise

The purpose of this exercise is to make a publication-ready plot using your `dplyr` and `ggplot2` skills. Open the journal article "Riddell_Annals_Hom-Sui-Disparities.pdf". Read the abstract and have a look at Figure 3. The data you need is "CDC_Males.csv". You'll have 1 hour to see what you can come up with in groups or individually!

You got this. Don't worry about making it exactly, try and see how far you can get. You're encouraged to work together if you want to and exchange tips/tricks you figured out. 

I'll leave these here:

* https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html
* http://colorbrewer2.org

Hint: You may need to use some `dplyr` skills from the first session

```{r}
# Replicate Figure 3
CDC_firearm_only <- read_csv("CDC_Males.csv") %>% 
  filter(
    type == "Firearm",
    !ST %in% c("US", "HI", "DC")
  )

corr <- cor.test(
  x = CDC_firearm_only$adjusted.homicide.White,
  y = CDC_firearm_only$adjusted.suicide.White,
  method = "spearman"
)

rho <- corr$estimate

ggplot(
  data = CDC_firearm_only, 
  aes(x = adjusted.suicide.White, y = adjusted.homicide.White)
) +
  geom_point(
    aes(fill = gun.house.prev.category, size = average.pop.white), 
    shape = 21
  ) + 
  scale_fill_manual(values = c("#fef0d9", "#fdcc8a", "#fc8d59", "#d7301f")) + 
  scale_size_area(
    breaks = c(200000, 500000, 1500000, 3000000, 7000000),
    labels = c("200k", "500k", "1.5m", "3m", "7m"), 
    max_size = 15
  ) +
  theme_bw() +
  labs(
    x = "White suicide rate (per 100,000 per year)",
    y = "White homicide rate (per 100,000 per year)"
  ) +
  geom_text_repel(aes(label = ST)) +
  geom_text(
    aes(x = 25, y = 0.75, label = paste("Spearman's rho:", round(rho, 2))), 
    check_overlap = TRUE) + 
  guides(
    fill = guide_legend(title = "Gun ownership", override.aes = list(size = 5), order = 1),
    size = guide_legend(title = "White population"), order = 2
  )
```

