---
title: "Tidy Data"
output:
  html_document:
    df_print: paged
---

```{r setup}
library(tidyverse)
library(gapminder)

# Toy data from https://github.com/jennybc/lotr-tidy

lotr <- tribble(
                         ~film,    ~race, ~female, ~male,
  "The Fellowship Of The Ring",    "Elf",   1229L,  971L,
  "The Fellowship Of The Ring", "Hobbit",     14L, 3644L,
  "The Fellowship Of The Ring",    "Man",      0L, 1995L,
              "The Two Towers",    "Elf",    331L,  513L,
              "The Two Towers", "Hobbit",      0L, 2463L,
              "The Two Towers",    "Man",    401L, 3589L,
      "The Return Of The King",    "Elf",    183L,  510L,
      "The Return Of The King", "Hobbit",      2L, 2673L,
      "The Return Of The King",    "Man",    268L, 2459L
  )


```

## Your Turn 1

Use `gather()` to reorganize `table4a` into three columns: **country**, **year**, and **cases**.

```{r}
gather(table4a, "year", "cases", -country)
```

## Your Turn 2

1. Use `spread()` to reorganize `table2` into four columns: country**, **year**, **cases**, and **population**. 
2. Create a new variable called `prevalence` that divides `cases` by `population` multiplied by 100000.
3. Pass the data frame to a ggplot. Make a scatter plot with `year` on the x axis and `prevalence` on the y axis. Set the color aesthetic (`aes()`) to `country`. Use `size = 2` for the points. Add a line geom. 

```{r}
table2 %>% 
  spread(type, count) %>% 
  mutate(prevalence = cases / population) %>% 
  ggplot(aes(year, prevalence, color = country)) + 
     geom_point(size = 2) +
     geom_line() + 
     scale_x_continuous(breaks = c(1999L, 2000L))
```

## Your Turn 3

Gather the 5th through 60th columns of `who` into a key column: value column pair named **codes** and **n**. Then select just the `country`, `year`, `codes` and `n` variables.

```{r}
who %>% 
  gather("codes", "n", 5:60) %>% 
  select(country, year, codes, n)
```

## Your Turn 4

Use the `cases` data below. Separate the `sex_age` column into **sex** and **age** columns.

```{r}
cases <- tribble(
   ~id,     ~sex_age,
   "1",    "male_56",
   "2",  "female_77",    
   "3",  "female_49"
)
separate(cases, sex_age, into = c("sex", "age"))
```

## Your Turn 5



## Your Turn 6: Challenge!

There are two CSV files in this folder containing SEER data in breast cancer incidence in white and black women. For both sets of data:

1. Import the data
2. Gather the columns into 2 new columns called `year` and `incidence`
3. Add a new variable called `race`. Remember that each data set corresponds to a single race.
4. Bind the data sets together using `bind_rows()` from the dplyr package. Either save it as a new object or pipe the result directly into the ggplot2 code.
5. Plot the data using the code below. Fill in the blanks to have `year` on the x-axis, `incidence` on the y-axis, and `race` as the color aesthetic. 

```{r}
breast_ca_black <- read_csv("breast_ca_black.csv")
breast_ca_white <- read_csv("breast_ca_white.csv")
breast_ca_black <- breast_ca_black %>% 
  gather("year", "incidence") %>% 
  mutate(race = "black")

breast_ca_white <- breast_ca_white %>% 
  gather("year", "incidence") %>% 
  mutate(race = "white")

breast_ca <- bind_rows(breast_ca_black, breast_ca_white)

breast_ca %>% 
  ggplot(aes(x = year, y = incidence, color = race, group = race)) + 
    geom_point() +
    geom_line() + 
    scale_x_discrete(breaks = seq(1975, 2011, by = 5)) # show years by fives
```

***

# Take Aways

Data comes in many formats but R prefers just one: _tidy data_.

A data set is tidy if and only if:

1. Every variable is in its own column
2. Every observation is in its own row
3. Every value is in its own cell (which follows from the above)

What is a variable and an observation may depend on your immediate goal.

