---
title: "Data validation in R and Python"
format: html
---

```{r}
#| label: setup
library(tidyverse)
library(pointblank)
worlds_fairs <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-08-13/worlds_fairs.csv')
```

## Your Turn 1: Discussion

What's the strangest thing you've ever seen in your data? What's are some times changes in data broke your code?

## Your Turn 2

Using the `worlds_fairs` data frame, use pointblank to verify that all start years are greater than or equal to 1851, the year of the first Worlds Fair:

1. Create an agent
2. Specify that the `start_year` column is greater than or equal to 1851
3. Interrogate the data

You can find the reference page for the many pointblank validation functions here: https://rstudio.github.io/pointblank/reference/index.html#validation-expectation-and-test-functions

```{r}
______(worlds_fairs) |> 
  _______(______, value = _____) |> 
  _____()
```

## Your Turn 3

Validate the following steps:

* `visitors`, `cost`, `area`, and `attending_countries` are all greater than or equal to 0. Allow for them to also be missing with `na_pass = TRUE`.
* The `start_date` is less than or equal to the `end_date`
* For expos of the category "World Expo", `start_date` is increasing

```{r}
worlds_fairs |> 
  mutate(
    start_date = my(paste0(start_month, "-", start_year)),
    end_date = my(paste0(end_month, "-", end_year))
  ) |> 
  create_agent(tbl_name = "worlds_fair") |> 
  # ... fill in the rest
  interrogate()
```

## Your Turn 4

Validate the following steps:

* The following variables are numeric:
  - start_month
  - start_year
  - end_month
  - end_year
  - area
  - visitors
  - cost
  - area
  - attending_countries
  
* The following variables are character:
  - name_of_exposition
  - country
  - city
  - category
  - theme
  - notables

```{r}
worlds_fairs |> 
  create_agent(tbl_name = "worlds_fair") |> 
  # ... fill in the rest |> 
  interrogate()
```

## Your Turn 5

Validate the following steps:

* All rows are distinct
* All rows have complete values of the start and end month and year, and the name of the exhibition

```{r}
worlds_fairs |> 
  create_agent(tbl_name = "worlds_fair") |> 
  # ... fill in the rest |> 
  interrogate()
```

## Your Turn 6

Run the following validation and inspect the report:

```{r}
al <- action_levels(
  warn_at = .01, 
  stop_at = .2
)

agent_report <- worlds_fairs |> 
  create_agent(tbl_name = "worlds_fair", actions = al) |> 
  col_vals_not_in_set(country, c("South Korea", "North Korea")) |> 
  col_vals_between(attending_countries, 10, 100) |> 
  rows_distinct() |> 
  rows_complete() |> 
  interrogate()

agent_report
```

Extract the failing data with `get_data_extracts()`. 
Identify why the set of rows failed given the expectation. 

```{r}

```

## Your Turn 7: Challenge!

For this exercise, let's use a new dataset

```{r}
english_monarchs_marriages <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-08-20/english_monarchs_marriages_df.csv')
```

Take a look at `english_monarchs_marriages` in detail.

```{r}
english_monarchs_marriages
```

We're going to clean this data frame up. First, though, write down your expectations of the data. What are the properties of the table, rows, columns, and values that you expect? 

Put these expectations down as code via pointblank and validating the data as-is. Use `get_sundered_data()` and `get_data_extracts()` to explore what parts of the data are passing and what parts are not.

```{r}

```

Now change the data so they pass your data validation specifications. 

```{r}

```


***

# Take aways

* Data is weird and can get weirder 
* Writing down expectations about data can help identify issues now and in the future
* pointblank offers an API for doing this:

```r
create_agent(.df) |> 
  # ... validation steps
  interrogate()
```

* With pointblank, we can assess properties of the data values with `col_vals_*()`, columns with `col_is_*()`, rows with `row_*()`, and at the table level with `*_match()`
* Investigate data validation results with `get_sundered_data()` and `get_data_extracts()`
