---
title: "Visualizing Data"
format:
  html:
    df-print: paged
---

```{r}
#| label: setup
library(tidyverse)
diabetes <- read_csv("diabetes.csv")
```

## Your Turn 10

1. Change the color scale by adding a scale layer. Experiment with `scale_color_distiller()` and `scale_color_viridis_c()`. Check the help pages for different palette options.
2. Set the color aesthetic to `gender`. Try `scale_color_brewer()`.
3. Set the colors manually with `scale_color_manual()`. Use `values = c("#E69F00", "#56B4E9")` in the function call.
4. Change the legend title for the color legend. Use the `name` argument in whatever scale function you're using.

```{r}
diabetes |> 
  ggplot(aes(waist, hip, color = weight)) + 
  geom_point()
```

## Your Turn 11

1. Change the theme using one of the built-in theme functions. 
2. Use `theme()` to change the legend to the bottom with `legend.position = "bottom"`.
3. Remove the axis ticks by setting the `axis.ticks` argument to `element_blank()`
4. Change the font size for the axis titles. Use `element_text()`. Check the help page if you don't know what option to change.

```{r}
diabetes |> 
  ggplot(aes(waist, hip, color = weight)) + 
  geom_point() 
```

## Your Turn 12

Take a look at the `diamonds` data set from ggplot2. How many rows does it have?

```{r}
diamonds
```

1. Make a scatterplot of `carat` vs. `price`. How's it look?
2. Try adjusting the transparency. 
3. Replace `geom_point()` with 2d bins.
4. Try hex bins.

```{r}

```

## Your Turn 13

1. Add a title. 
2. Change the x and y axis labels to include the units (inches for `hip` and pounds for `weight`). You can use either `labs()` or `xlab()` and `ylab()`
3. Add `scale_linetype()` and set the `name` argument to "Sex".

```{r}
ggplot(diabetes, aes(weight, hip, linetype = gender)) +
  geom_jitter(alpha = .2, size = 2.5) +
  geom_smooth(color = "black", se = FALSE) + 
  theme_bw(base_size = 12)
```


## Your Turn 14

Save the last plot and then locate it in the files pane.

```{r}

```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**. You may not even need to add scales and themes!

```{r}
#| eval: false
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>)) +
  <SCALE_FUNCTION>() + 
  <THEME_FUNCTION>()
```

