---
title: "Whole Game: Analyzing the diabetes data set"
output: html_document
bibliography: references.bib
---

# Introduction

In this exercise, we'll look at a report generated completely in R using R Markdown.
We're using an intentionally simplified report: what we're telling the readers, how we fit models, and so on, are all much simpler than you would use in practice.
Instead, focus on the big idea here: reproducible reports!

```{r setup, include = FALSE}
library(tidyverse)
library(gtsummary)
```

```{r prep-data, include = FALSE}
diabetes <- read_csv("diabetes-buckingham.csv") %>% 
  mutate(diabetic = ifelse(glyhb >= 6.5, "Diabetic", "Healthy"))

description <- list()
```

# Methods

These data were collected as part of a study to estimate the prevalence of obesity, diabetes, and other cardiovascular risk factors in central Virginia for African-Americans [@Willems1997].
We sought to assess the relationship between waist-to-hip ratio on hemoglobin A1c and diabetes diagnosis.
In this truncated report, we include basic descriptive statistics on participants by diabetes status.
We also fit two regression models: 1) an ordinary least-squares regression model of the impact of waist-to-hip ratio on hemoglobin A1c, and 2) a logistic regression model of the impact of waist-to-hip ratio on diabetes status.
We adjusted both models for age.

# Results

```{r descriptive-stats, eval = TRUE, echo = FALSE}
description <- diabetes %>% 
  # calculate basic statistics by diabetes status
  group_by(diabetic) %>% 
  summarise(n = n(), across(c(glyhb, ratio, age), mean, na.rm = TRUE)) %>% 
  # clean up data to include in text
  mutate(
    across(where(is.numeric), round),
    diabetic = fct_explicit_na(diabetic, na_level = "Missing")
  ) %>% 
  # split into a list of data frames by diabetes status
  split(.$diabetic)
```

There were `r description$Diabetic$n` participants with diabetes, who had an average A1c of `r description$Diabetic$glyhb`, and `r description$Healthy$n` participants with diabetes, who had an average A1c of `r description$Healthy$glyhb`.
Non-diabetic participants were, on average, younger than those with diabetes (`r description$Healthy$age` vs `r description$Diabetic$age`), There were `r description$Missing$n` participants who were missing data.
These participants were also younger, on average, than those with diabetes (`r description$Diabetic$age` vs `r description$Missing$age`) See Table 1.

```{r table-one, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
lbls <- list(
  diabetic ~ "Diabetes Status", 
  glyhb ~ "Hemoglobin A1c",
  ratio ~ "Waist/Hip Ratio", 
  age ~ "Age", 
  chol ~ "Cholestorol", 
  gender ~ "Sex", 
  frame ~ "Body Frame"
)

diabetes %>% 
  select(diabetic, glyhb, ratio, age, chol, gender, frame) %>% 
  filter(!is.na(diabetic)) %>% 
  tbl_summary(by = diabetic, label = lbls, missing_text = "(Missing)") %>% 
  add_difference()
```

```{r figure-one, eval = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
library(_______)

fig1a <- diabetes %>% 
  drop_na() %>% 
  ggplot(aes(ratio, glyhb)) + 
  geom_point(shape = 21, fill = "grey80", color = "white", size = 2) + 
  geom_smooth(size = 1, color = "steelblue", se = FALSE, method = "lm") + 
  scale_x_log10() +
  theme_minimal(14) +
  labs(
    x = "Hip/waist ratio",
    y = "Hemoglobin A1c",
    tag = "A"
  )

fig1b <- diabetes %>% 
  drop_na() %>% 
  mutate(diabetic = factor(diabetic, levels = c("Healthy", "Diabetic"))) %>% 
ggplot(aes(ratio, fill = diabetic)) + 
  geom_density(color = "white", alpha = .8) +
  theme_minimal(14) +
  theme(
    legend.title = element_blank(),
    legend.position = "bottom"
  ) +
  labs(x = "Hip/waist ratio", tag = "B") +
  scale_fill_manual(values = c("steelblue", "firebrick"))

_______
```

In Figure \_\_\_\_\_\_\_a, we show the unadjusted relationship between A1c and waist-hip ratio (WH ratio is shown on the log scale).
As WH ratio increases, A1c also increases.
In Figure \_\_\_\_\_\_\_b, we show the density of WH ratio among those with and without diabetes.
Participants with diabetes had a higher WH ratio than those without diabetes.

```{r table-two, eval = FALSE, echo = FALSE}
lbls_reg <- list(age ~ "Age", ratio ~ "Waist/Hip Ratio")

diabetes <- diabetes %>% 
  drop_na() %>% 
  mutate(diabetic = factor(diabetic, levels = c("Healthy", "Diabetic")))

linear_mod_tbl <- lm(glyhb ~ ratio + age, data = diabetes) %>% 
  _______(label = lbls_reg)

logistic_mod_tbl <- glm(factor(diabetic) ~ ratio + age, data = diabetes, family = binomial()) %>% 
  _______(label = lbls_reg, exponentiate = TRUE)

_______(
  tbls = list(linear_mod_tbl, logistic_mod_tbl),
  tab_spanner = c("**Linear model (HbA1c)**", "**Logistic model (Diabetes status)**")
)
```

In Table 2, we present two regression models: linear and logistic.
In the linear model, we analyze the association of WH ratio an A1c adjusted for age.
In the logistic model, we analyze the association of WH ratio on diabetes status adjusted for age.
Increased WH ratio was associated with both higher A1c and a higher likelihood of being diabetic.

# Discussion

In this R Markdown file, we generated a polished report of the diabetes dataset.
Amazingly, when we were provided with an updated dataset, everything updated perfectly!

# References
