---
title: "Manage code pipelines with targets"
format: html
---

```{r}
#| label: setup
#| include: false
library(targets)
# this loads the `copy_targets_template()` helper function we'll use in this exercise
source("copy_targets_template.R")
```

## Your Turn 1

1. Use `tar_script()` in the console, then open `_targets.R` (you can also use `tar_edit()` in the console to open it for you). Read the resulting script.
2. Predict how many targets there are, what they are called, and any other dependencies in your code. Run `tar_visnetwork()` in the console to check if you were right.
3. Run `tar_make()` in the console, then run `tar_visnetwork()` again. What's different? Try running `tar_make()` again.

## Your Turn 2

1. Run this code to make a fresh copy of `_targets.R`:

```{r}
# this function is defined in the `setup` chunk above!
copy_targets_template("your_turn_2.R")
```

2. Add targets to the `tar_plan()` at the bottom of `_targets.R` for reading in the diabetes dataset:
  * Tell targets to monitor the file with `tar_file(diabetes_file, "diabetes-buckingham.csv")`.
  * Read the file in by adding this target (fill in the blank with the name of the file path target!):

```{r}
#| label: diabetes-target
#| eval: false
diabetes = read_csv(__________, col_types = cols()) |> 
  mutate(diabetic = ifelse(glyhb >= 6.5, "Diabetic", "Healthy"))
```

3. Predict what you think the dependency graph will look like, then run `tar_visnetwork()` in the console
4. Run `tar_make()` in the console
5. Take a look at `diabetes` with `tar_read()`

## Your Turn 3

Something not quite working from the last exercise? Run this code to update `_targets.R` and catch up:

```{r}
copy_targets_template("your_turn_3.R")
```

1. Change `diabetes_file` to use `diabetes.csv` instead
2. Run `tar_outdated()` in the console. What's this telling you? Confirm with `tar_visnetwork()`
3. Predict which targets are going to re-run, then run `tar_make()`. Were you right?
4. Confirm that `diabetes` has changed by looking at it with `tar_read()`. The new dataset should have 403 rows and 20 columns.

## Your Turn 4

Something not quite working from the last exercise? Run this code to update `_targets.R` and catch up:

```{r}
copy_targets_template("your_turn_4.R")
```

1. Add `source("R/functions.R")` to `_targets.R`
2. Add "gtsummary" to the `packages` argument of `tar_option_set()`
3. Create a new target called `table_one` using `create_table_one(diabetes)`
4. Run `tar_visnetwork()` and `tar_make()`
5. Take a look at the target you just created

## Your Turn 5

Something not quite working from the last exercise? Run this code to update `_targets.R` and catch up:

```{r}
copy_targets_template("your_turn_5.R")
```

1. Open `R/functions.R` and modify `create_table_one()`: Add the argument `missing_text = "(Missing)"` to `tbl_summary()`. Make sure to save your file after you've made the change.
2. Run `tar_outdated()` in the console, then look at `tar_visnetwork()`
3. Predict which targets are going to re-run, then run `tar_make()`. Were you right?

## Your Turn 6

1. Let's skip ahead in our analysis. Run this code to create a new `_targets.R` file that has most of the targets we want to create. Open the file to make sure it was created correctly. 

```{r}
#| label: create-targets-r
copy_targets_template("your_turn_6.R")
```

2. Run `tar_make()` to catch up on the new targets.
3. Open `report.qmd`. Which targets does this report depend on? Look for `tar_load()` and `tar_read()`.
4. Add a new target to generate your report. The format for for rendering a Quarto file is `tar_quarto(name_of_target, "name_of_report.qmd")`.
5.  Run `tar_visnetwork()` and `tar_make()`. Open your rendered report from the Files pane.

## Your Turn 7

Something not quite working from the last exercise? Run this code to update `_targets.R` and catch up:

```{r}
copy_targets_template("your_turn_7.R")
```

Confirm that you can reproduce your entire pipeline from scratch. In the console:

1.  Run `tar_destroy()`
2.  Run `tar_make()`

# Take aways

- targets allows you to write a clear description of your analysis pipeline using function-based declarations
- targets knows which parts of your code depend on the others, so it only re-runs what it needs to
- `_targets.R` is how you tell targets what to run: end in a `list()` of `tar_target()`s or use `tarchetypes::tar_plan()` to use the format `target = action(other_target)`.
- `tar_make()` renders your pipeline, and `tar_visnetwork()` visualizes it
- Render Quarto reports in targets with `tarcheypes::tar_quarto()`

