---
title: "R Markdown: Whole Game"
output: html_document
---

```{r setup}
library(tidyverse)

copy_rmd_template <- function(template) {
  if (template == "your_turn_3.Rmd") {
    file.copy(
      file.path("templates", "references.bib"),
      "references.bib",
      overwrite = TRUE
    )
  }
  copied <- file.copy(
    file.path("templates", template),
    "whole_game.Rmd",
    overwrite = TRUE
  )
  
  if (!copied) {
    stop(
      file.path("templates", template), " not found. ", 
      "Are you sure you're in the right working directory?",
      call. = FALSE
    )
  }
  
  template_r <- usethis::ui_path(file.path("templates", template))
  targets_r <- usethis::ui_path("whole_game.Rmd")
  usethis::ui_done("Copying {template_r} to {targets_r}")
}
```

*For this set of exercises, we'll work through an R Markdown report in `whole_game.Rmd`. Open up that file and follow along with these exercises:*

## *Your Turn 1*

1.  *Turn on Visual R Markdown mode for `exercises.Rmd` (this document) AND `whole_game.Rmd`*
2.  *In your local R session, run the code in the code chunk called "prep-data". Look at the chunk options (everything after "r prep-data". What do you think setting the `include` option does?*
3.  *Knit the document and take a look. Were you right?*

## *Your Turn 2*

1.  *Add a citation to the methods section for the data. Use the citation wizard to search PubMed for an article called "Prevalence of coronary heart disease risk factors among rural blacks: A community-based study"*

## *Your Turn 3*

*Something not quite working from the last exercise? Run this code to update `whole_game.Rmd` and catch up*

```{r, eval=FALSE, echo = FALSE}
copy_rmd_template("your_turn_3.Rmd")
```

1.  *Take a look at the chunk called "descriptive-stats." What do you think `echo = FALSE` does, and how is it different from `include = FALSE`? What do you think `eval = FALSE` means?*
2.  *Read the code in "descriptive-stats". What are we preparing?*
3.  *Look at the narrative text right after "descriptive-stats". It's full of bits of code. What are we doing here, and what do you think the output should look like?*
4.  *Change `eval` to `TRUE` in the "descriptive-stats" code chunk and re-knit. Look at this part of the document. What does the output look like?*

## *Your Turn 4*

*Something not quite working from the last exercise? Run this code to update `whole_game.Rmd` and catch up*

```{r, eval=FALSE, echo = FALSE}
copy_rmd_template("your_turn_4.Rmd")
```

1.  *In the code chunk called "table-one", create a descriptive table with `tbl_summary()`*
2.  *Pipe the results of `tbl_summary()` to `add_difference()`.*
3.  *Change `eval` to `TRUE` and re-knit*

## *Your Turn 5*

*Something not quite working from the last exercise? Run this code to update `whole_game.Rmd` and catch up*

```{r, eval=FALSE, echo = FALSE}
copy_rmd_template("your_turn_5.Rmd")
```

1.  *In the code chunk called "figure-one", load the patchwork library*

2.  *At the end of the code chunk, include this code: `fig1a + fig1b` . patchwork allows you to join ggplots with `+` (and more).*

3.  *At the top of the document, change `output: html_document` to this code (be mindful to keep the indents as is):*

``` {.yaml}
output: 
bookdown::html_document2:
number_sections: false
```

4.  *In "figure-one", add a chunk option: `fig.cap = "A1c and diabetes status by WH ratio"` . Then, where it says "Figure \_\_\_\_\_", add* \@ref(fig:figure-one)

5.  *In "figure-one", change `eval` to `TRUE` and re-knit*

## *Your Turn 6*

*Something not quite working from the last exercise? Run this code to update `whole_game.Rmd` and catch up*

```{r, eval=FALSE, echo = FALSE}
copy_rmd_template("your_turn_6.Rmd")
```

1.  *In the code chunk called "table-two", pass the results of both `lm()` and `glm()` to `tbl_regression()`*
2.  *Merge the two model tables using `tbl_merge()`*
3.  *change `eval` to `TRUE` and re-knit*

***Stretch goal**: The description for Table 2 is pretty sparse. It would probably be better to have some numbers included in the narrative text. Read gtsummary's [vignette on inline text](http://www.danieldsjoberg.com/gtsummary/articles/inline_text.html) and add a few important details.*

## *Your Turn 7*

*Something not quite working from the last exercise? Run this code to update `whole_game.Rmd` and catch up*

```{r, eval=FALSE, echo = FALSE}
copy_rmd_template("your_turn_7.Rmd")
```

*Previously, we only had data for one location in the study---Buckingham. We've just received data from the other site, Louisa, and we need to update the tables, figures, and narrative text...*

*...but since we used R Markdown, we can just load the new data set, and R will update everything for us.*

1.  *Change `read_csv()` to load `diabetes.csv` instead of `diabetes-buckingham.csv`*
2.  *Add the `location` variable to Table 1*
3.  *Re-knit. How's it look?*
