---
title: "Table Recreation Challenge: R"
format: html
---

## Your Turn: Table Recreation Challenge

Recreate the table shown in `recreation.png` using the data and labels below.

### Hints

You will need these functions not covered in the slides:

- `fmt_markdown()`: render HTML markup in a column
- `fmt_percent()`: format values as percentages
- `fmt_currency()`: format values as currency
- `cols_align()`: set column alignment
- The `rows` argument in formatting functions lets you apply formatting to specific rows
- `column_labels.font.weight` in `tab_options()` controls label boldness

You may like to consult the gt reference: https://gt.rstudio.com/reference/index.html

```{r}
library(gt)

reg_df <- data.frame(
  Method = c(
    "Average",
    "Covariance",
    "Net Compensation<sup>&dagger;</sup>",
    "Weighted Average<sup>&Dagger;</sup>",
    "Mean Residual<br>Difference<sup>&oplus;</sup>",
    "OLS"
  ),
  r2 = c(12.4, 12.4, 12.5, 12.6, 12.8, 12.9),
  ratio_g = c(0.996, 0.996, 0.980, 0.964, 0.895, 0.837),
  ratio_gc = c(1.001, 1.001, 1.006, 1.011, 1.032, 1.050),
  comp_g = c(-46, -46, -232, -411, -1208, -1872),
  comp_gc = c(4, 4, 34, 62, 188, 293),
  mrd = c(-50, -50, -266, -473, -1396, -2165),
  fair_cov = c(6, 6, 31, 56, 164, 256)
)

r2_label <- html("R<sup>2</sup>")
g_label <- html("<em>g</em>")
gc_label <- html("<em>g</em><sup>c</sup>")
mrd_label <- html("Mean<br>Residual<br>Difference")
fair_cov_label <- html("Fair<br>Covariance")

ratio_spanner <- html("Predictive<br>Ratio")
comp_spanner <- html("Net<br>Compensation")

source_note_text <- html(paste0(
  "&dagger; &lambda; = 10000, &Dagger; &alpha; = 0.2, &oplus; &lambda; = 30000<br>",
  "<em>Note:</em> Measures calculated based on cross-validated predicted values and ",
  "sorted on net compensation. Best performing hyperparameters for each ",
  "estimator (with respect to fairness measures) are displayed. Performance ",
  "for covariance method was the same for all <em>m</em>. <em>g</em><sup>c</sup> is the complement of <em>g</em>."
))
```

```{r}
# Recreate the table from `recreation.png`

```
