---
title: "Tables in R Markdown"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse)
library(gapminder)
library(gtsummary)
library(broom)
library(knitr)
```

## Your Turn 1

1. Using Visual R Markdown, create a markdown table that represents these data: 
2. Knit

## Your Turn 2

1. Using the `tidy()` function from broom, turn `response_model` and `marker_model` into dataframes.
2. Bind the two resulting model dataframes on top of each other to create one dataframe
3. Turn the resulting dataframe into a table using `kable()`
4. Knit

```{r}
response_model <- glm(response ~ age + stage, data = trial, family = binomial)
marker_model <- lm(marker ~ trt + stage + age, data = trial)


```

## Your Turn 3

1. Create a contingency table of `trial` (a dataset from gtsummary) using `tbl_cross()`: set `trt` to the rows and `response` to the columns.
2. Add a p-value
3. Knit

```{r}

```

## Your Turn 4

## Your Turn 5

## Your Turn 6


# Take aways
- 
